Tables

### 2023-11-16 output/paper/figures.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

---
title: "Figures"
format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers
link-citations: true
linkcolor: blue
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

num_digits_est = 3
num_digits_sig = 2
```

- Use log-scale when plotting relative measures of effects.
- Figure titles and legends should be submitted as single paragraph.

>>>>

---
title: "Figures"

format:
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

link-citations: true
linkcolor: blue

editor_options: 
  chunk_output_type: console

execute:
  echo: false
  warning: false
  error: false
  message: false
---

```{r}
here::i_am("output/paper/figures.qmd")
```

```{r vars}
source(here::here("code/dictionaries.R"))
source(here::here("code/data.R"))
source(here::here("code/proc_res.R"))

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params <- params(is_hpc = is_hpc)
threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params$steps$rq2$preproc_outcome$missings$k
```

- Figure titles and legends should be submitted as single paragraph.

### 2023-11-16 output/paper/manuscript.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

format:
  docx:
    reference-doc: reference_document.docx
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    keep-tex: true
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

@@

@@

::: {#refs}
:::

>>>>

format:
  docx:
    reference-doc: reference_document.docx
  pdf:
    pdf-engine: lualatex
    number-sections: true
    toc: false
    keep-tex: false
    include-in-header:
      text: |
        \usepackage{lineno}
        \usepackage{rotating}
        \usepackage{float}
        \usepackage{typearea}
    include-before-body:
      text: |
        \linenumbers

@@

```{r}
here::i_am("output/paper/manuscript.qmd")
```

@@

### 2023-11-16 output/paper/sections/methods.qmd
Added table LLOQs for steroids. Started kinda intro to paper.
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

@@

```{r params-preproc}
threshold_within_lod <- 20
threshold_overall_lod <- 10
threshold_within_nan <- 40
threshold_overall_nan <- 30
knn_k <- 5
```

@@

Concentrations of the corticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected. For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall. We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall.
The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall. We used `r knn_k` nearest neighbors. We natural log-transformed them to improve model fit, assessed with posterior predictive checks. To do so, replicated data were simulated with the fitted models and compared to the observed data. We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021]. Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.

Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed. Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd]. Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study. Remaining missing values were imputed similarly using kNN. Values of the chemicals were expressed in $\mu$grams per litre.

@@

To estimate the average contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023]. The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately. The average contrasts were computed using the estimated balancing weights above. Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].

>>>>

```{r vars}
source(here::here("code/dictionaries.R"))

params <- params(is_hpc = FALSE)
threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params$steps$rq2$preproc_outcome$missings$k
```

@@

@@

Concentrations of the corticosteroids were classified as quantifiable, below the \acr{loq}, possible interference or out of range, and not detected. For each metabolite, we computed the fraction of values below the \acr{loq} and not detected, both within each cohort and overall. We proceeded to impute these values using half the value of the corresponding \acr{lloq}, for those metabolites that had less than `r threshold_within_lod`% of missings within each cohort and `r threshold_overall_lod`% of missings overall. Information about the \acr{lloq} for the corticosteroids is provided in @tbl-lloq-mets. The remaining missing values were imputed using kNN from the `VIM` R package [@KowarikTempl:2016], for those metabolites that had less than `r threshold_within_nan`% of remaining missings within each cohort and `r threshold_overall_nan`% of remaining missings overall. We used `r knn_k` nearest neighbors. We natural log-transformed them to improve model fit, assessed with posterior predictive checks. To do so, replicated data were simulated with the fitted models and compared to the observed data. We used the `check_predictions` function from the `performance` R package using the default arguments [@LudeckeBen-ShacharPatil:2021]. Values of cortisol production and cortisone production were expressed in nanograms per millilitre, whereas values of cortisol metabolism and 11bHSD activity were unitless.

Concentrations of the non-persistent \acr{edcs} were classified as quantifiable, below the \acr{lod}, possible interference or out of range, and not analysed. Concentrations below the \acr{lod} were singly imputed using a quantile regression approach for the imputation of left-censored missing data, as implemented in the `impute.QRILC` function from the `imputeLCMD` R package [@lazar2015imputelcmd]. Information about the lower limits of detection can be found in @HaugSakhiCequier:2018. Chemicals with more than 70% of observations below the \acr{lod} were not considered in the present study. Remaining missing values were imputed similarly using kNN. Values of the chemicals were expressed in $\mu$grams per litre.

@@

To estimate the average contrasts, we used the `avg_comparisons` function from the `marginaleffects` R package [@Arel-Bundock:2023]. The two counterfactual datasets were obtained by setting the exposures levels to 90th percentile ($d_1$) and the 10th percentile ($d_2$), for each cohort separately. The average contrasts were computed using the estimated balancing weights above. Robust standard errors were computed with the `sandwich` R package, using cohort as variable indicating clustering of observations [@Zeileis:2004; @ZeileisKollGraham:2020].
For each outcome, we report the results as differences between average contrasts.

### 2023-11-16 output/paper/si.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

```{r}
source("code/dictionaries.R")
source("code/data.R")
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
```

@@

mets_info <- readODS::read_ods(
  path = "docs/steroids.ods"
) |>
  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
gt::gt(mets_info,
       groupname_col = "type") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    analyte = "Metabolite",
    acronym = "Symbol",
    hmdb_id = "HMDB ID",
    cas_number = "CAS number"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  ) |>
  gt::tab_footnote(
    footnote = "Human Metabolome Database",
    locations = gt::cells_column_labels(
      columns = hmdb_id
    )
  ) |>
  gt::tab_footnote(
    footnote = "Chemical Abstracts Service",
    locations = gt::cells_column_labels(
      columns = cas_number
    )
  )
```

>>>>

```{r}
here::i_am("output/paper/si.qmd")
```

```{r vars}
source(here::here("code/dictionaries.R"))
source(here::here("code/data.R"))
source(here::here("code/proc_res.R"))

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params_ <- params(is_hpc = is_hpc)
threshold_within_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params_$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params_$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params_$steps$rq2$preproc_outcome$missings$k
```

@@

mets_info <- readODS::read_ods(
  path = here::here("docs", "steroids.ods")
) |>
  tidylog::select(analyte, acronym, type, hmdb_id, cas_number)
gt::gt(mets_info,
       groupname_col = "type") |>
  gt::sub_values(
    values = c("NA"),
    replacement = ""
  ) |>
  gt::cols_label(
    analyte = "Metabolite",
    acronym = "Symbol",
    hmdb_id = "HMDB ID",
    cas_number = "CAS number"
  ) |>
  gt::tab_style(
    locations = gt::cells_row_groups(group = dplyr::everything()),
    style = list(gt::cell_text(weight = "bold"))
  ) |>
  gt::tab_footnote(
    footnote = "Human Metabolome Database",
    locations = gt::cells_column_labels(
      columns = hmdb_id
    )
  ) |>
  gt::tab_footnote(
    footnote = "Chemical Abstracts Service",
    locations = gt::cells_column_labels(
      columns = cas_number
    )
  )
```

### 2023-11-16 output/paper/tables.qmd
Modified how paper is rendered, moving away from targets for that and simply calling quarto render on SI, tables, figures, and manuscript, and including sections of manuscript in manuscript.qmd. Still, cross ref does not work like this since they are rendered independently.

```{r}
source("code/proc_res.R")
```

```{r vars}
store <- ifelse(
  Sys.getenv("is_hpc"), 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

num_digits_est = 3
num_digits_sig = 2
```

>>>>

```{r}
here::i_am("output/paper/tables.qmd")
```

```{r vars}
source(here::here("code/dictionaries.R"))
source(here::here("code/data.R"))
source(here::here("code/proc_res.R"))

is_hpc <- FALSE
store <- ifelse(
  is_hpc, 
  paste0(
    "/PROJECTES/HELIX_OMICS/DATA_PREVIOUS_MIGRATION/",
    "lorenzoF/data/data_paper3/_targets/_targets_res"
  ), 
  "~/mounts/rstudioserver/data/data_paper3/_targets/_targets_res"
)

id_var <- "HelixID"
num_digits_est = 3
num_digits_sig = 2
params <- params(is_hpc = is_hpc)
threshold_within_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_within
threshold_overall_lod <- params$steps$rq2$preproc_outcome$llodq$threshold_overall
threshold_within_nan <- params$steps$rq2$preproc_outcome$missings$threshold_within
threshold_overall_nan <- params$steps$rq2$preproc_outcome$missings$threshold_overall
knn_k <- params$steps$rq2$preproc_outcome$missings$k
```
